Evaluating a simplified forecast model in comparison to the European forecasting hub ensemble

### 2022-09-27 paper/paper.Rmd
improve handling of forecast visualisation and close #14

---
title: Evaluating a simplified forecast model in comparison to the European forecasting hub ensemble
preprint: true
author: 
  - name: Sam Abbott
    affiliation: 1
    corresponding: true
    email: sam.abbott@lshtm.ac.uk
  - name: Katharine Sherratt
    affiliation: 1
  - name: Nikos Bosse
    affiliation: 1
  - name: Hugo Gruson
    affiliation: 1
  - name:  Johannes Bracher
    affiliation: 2
  - name: Sebastian Funk
    affiliation: 1
affiliation:
  - code: 1
    address: London School of Hygiene and Tropical Medicine, London, UK
  - code: 2
    address: Chair of Statistics and Econometrics, Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany
abstract: > 
  Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts, especially during the COVID-19 pandemic. However, these forecasts require a coordinated team to produce and are effectively black boxes. In other fields, resource usage has been reduced by training simplified models that reproduce some of the observed behaviour of more complex models. 
  Here we use observations of the behaviour of the European COVID-19 Forecast Hub ensemble combined with our own forecasting experience, to specify a series of model assumptions modelling incident reported cases of COVID-19. We then develop a forecast model using these assumptions and compare its performance to the European forecast hub ensemble in real-time from the 15th of January 2022 to the 19th of July 2022 over a four-week forecast horizon.
  We found that this surrogate model performed visibly relatively similarly to a subset of example locations though with increased uncertainty. Unsurprisingly, using the weighted interval score, we concluded that average performance was substantially worse relative to the ensemble forecast but that median performance was more comparable. The majority of the difference in performance was linked to a subset of forecasts with worse performance early in 2022. We note that despite performing worse than the hub ensemble the majority of the time our surrogate model showed better calibration, and was less liable to underpredict. The central part of the surrogate model’s forecast distribution performed more similarly to the ensemble forecast’s distribution but maintained a right skewed performance distribution.
  We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble. However, more work needs to be done to understand the implicit epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders' requirements. We also conclude that our findings may be largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements. Our work is potentially useful for forecast consumers as a tool for understanding the inherent assumptions of the forecasts they are making use of and to researchers aiming to develop forecast approaches that perform similarly, or better than, current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
bibliography: library.bib
output:
  bookdown::pdf_book:
    base_format: rticles::peerj_article
    extra_dependencies: ["float"]
---

@@

We evaluate forecasts for all locations and horizons quantitatively using the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], which is a quantile-based proper scoring rule that approximates the continuous ranked probability score (CRPS).
Both the WIS and CRPS are generalisations of the absolute error to evaluate probabilistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the European forecast hub [@Sherratt2022-ue].
We present WIS for the subset of forecasts we explore visually for both the ensemble and surrogate model by date and forecast horizon (1 and 4 weeks).

To understand the relative performance of the surrogate model compared to the ensemble model, we calculate the relative performance (rWIS) by dividing the WIS for the surrogate model by the WIS of the ensemble model for all locations and forecast horizons.
To maintain the propriety of this score, we do this after first taking the means of scores for the relevant stratification.
We explore relative performance by forecast horizon, by month and horizon, and by location and horizon.

In addition to presenting the WIS for a subset of locations and the relative WIS for all locations, we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% prediction intervals and by quantile. 
Lastly, we calculate and visualise the relative weighted interval score by quantile, stratified by forecast horizon.

@@

In our example set of locations, the absolute performance of the ensemble and the surrogate model was visually similar on the log scale in all locations at short forecast horizons though this varied by location.
Performance was most similar in Slovakia but in the United Kingdom and Germany in particular the surrogate model performed visibly worse for some forecast dates (Figure \@ref(fig:vis-forecasts)).
For both the ensemble and the surrogate, performance visibly decreased as the forecast horizon increased though the relative difference in performance between models was relatively stable across horizons.
In general, in the study period, the ensemble appeared to be liable to underpredict and the surrogate model appeared more likely to over-predict. 
Both models forecast large reductions in incidence in Poland during May that did not occur whilst only the ensemble forecast spuriously forecast similar large reductions in Germany during June.
In comparison to the ensemble model the surrogate model appeared less likely to place weight on unfeasibly large reductions in incidence during periods of declining incidence but on other hand was more likely to forecast continuing increases in incidence (for example in February in Slovakia and Poland).
Italy and Germany, which had the highest number of submitted models of the example locations had little clear difference in performance between the ensemble and surrogate models compared to other example locations.

@@

Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the forecast hub ensemble on the log scale though with greater uncertainty.
In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased toward underprediction, and the surrogate model appeared slightly biased toward overprediction.
Evaluating the relative performance of the surrogate model compared to the European forecast hub ensemble we found that the mean performance was substantially worse, which was driven by a relatively small subset of vey poorly performing forecasts.
Median forecast performance of the surrogate model was also worse when compared to forecasts from the ensemble though the majority of surrogate forecasts were within 50% of the performance observed for the ensemble forecast.
Performance varied by location and forecast date with the surrogate model performing worse in the first part of 2022 which may have been linked to incidence rates peaking across forecast locations linked to the spread of BA.2.
In general, the relative performance of the surrogate model degraded as forecast horizons increas with the distribution of relative performance having an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison to the hub ensemble.
The forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and bigger prediction intervals, compared to the surrogate model though the surrogate model tended to be overly uncertain at large intervals.
Evaluating point forecast performance indicated a similar pattern of performance as that observed using the full predictive distribution though relative performance of the surrogate model generally improved.

@@

Whilst we derived our surrogate model from a range of assumptions based on observing ensemble forecasts behaviour and the behaviour and structure of submitted models this was by no means as structured or complete a process as it could have been.
In follow-up work, a more rigorous approach to this could be taken to further refine this set of assumptions, in particular using the input of a wider pool of researchers. 
The findings from our study may also be useful for informing this improved set of assumptions. 
In addition, the model we derived based on our assumptions was likely not optimal both in terms of compute time and accuracy at reproducing ensemble-like behaviour.
Models with a more complex auto-correlation structure and more refined approaches to localised trends should be explored to improve relative performance to ensemble forecasts.
An example of a family of possible approaches are structural time series models which have many of the characteristics implied by our assumptions for how forecast ensembles typically operate.
As we identified that the tails of our predictive distributions were responsible for a large proportion of the difference in performance compared to the forecast ensemble it may be the case that post-processing of forecasts from our surrogate model would enhance their similarity to the forecast ensemble.
This seems likely to improve out-of-sample performance but does not help with the secondary aim of understanding the implicit assumptions driving the performance of multi-model, multi-team infectious disease forecast ensembles.
As we have hypothesised that the use of absolute scoring measures is inappropriate and leads to performance characteristics that are unlikely to be favoured by forecast stakeholders more work should be done in this area.
If new forecast ensemble methods are adopted as best practice by forecast hubs then follow-up work attempting to create surrogate forecast models should also use these approaches and this will likely alter the observed characteristics of the hub ensemble forecasts, for example, the tendency to be poorly calibrated.
In September 2022, GitHub announced support for hosted GitHub Action runners with additional compute power [@GitHubActionsLargeRunners]. Whilst a paid feature this may allow more compute attensive models, with fewer potental performance trade-offs, to be easily democratised though only if funds are available to support the hosting costs.
One potential research area is to explore forecasting methods that can be used with a range of compute resources though this would require extensive evaluation and documentation to make it clear to users what the trade-offs between compute usage and forecast performance are.
More work is needed to understand the best practice treatment of data revisions when evaluating forecasts and the potential sources for bias these may cause.
Lastly, here we have only explored a surrogate for an ensemble for a single disease, a limited set of locations, and a single target (incident cases), meaning our findings are difficult to generalise.
Follow-up work should explore whether this behaviour holds across diseases, locations, and epidemiological targets where the behaviour of ensembles is notably different [@REF].
However, this work is limited by the lack of large-scale forecast ensembling projects despite them showing obvious promise to improve the forecasts available to stakeholders.

>>>>

---
title: Evaluating a simplified forecast model in comparison to the European forecasting hub ensemble
preprint: true
author: 
  - name: Sam Abbott
    affiliation: 1
    corresponding: true
    email: sam.abbott@lshtm.ac.uk
  - name: Katharine Sherratt
    affiliation: 1
  - name: Nikos Bosse
    affiliation: 1
  - name: Hugo Gruson
    affiliation: 1
  - name:  Johannes Bracher
    affiliation: 2
  - name: Sebastian Funk
    affiliation: 1
affiliation:
  - code: 1
    address: London School of Hygiene and Tropical Medicine, London, UK
  - code: 2
    address: Chair of Statistics and Econometrics, Karlsruhe Institute of Technology (KIT), Karlsruhe, Germany
abstract: > 
  Multi-model and multi-team ensembles have become increasingly popular as an approach to increase the robustness and performance of infectious disease forecasts, especially during the COVID-19 pandemic. However, these forecasts require a coordinated team to produce and are effectively black boxes. In other fields, resource usage has been reduced by training simplified models that reproduce some of the observed behaviour of more complex models. 
  Here we use observations of the behaviour of the European COVID-19 Forecast Hub ensemble combined with our own forecasting experience, to specify a series of model assumptions modelling incident reported cases of COVID-19. We then develop a forecast model using these assumptions and compare its performance to the European forecast hub ensemble in real-time from the 15th of January 2022 to the 19th of July 2022 over a four-week forecast horizon using the weighted interval score.
  We found that this surrogate model performed visibly relatively similarly to a subset of example locations though with increased uncertainty. Unsurprisingly, we concluded that average performance was substantially worse relative to the ensemble forecast but that median performance was more comparable. The majority of the difference in performance was linked to a subset of forecasts with worse performance early in 2022. We note that despite performing worse than the hub ensemble the majority of the time our surrogate model showed better calibration, and was less liable to underpredict. The central part of the surrogate model’s forecast distribution performed more similarly to the ensemble forecast’s distribution but maintained a right skewed performance distribution.
  We conclude that our simplified forecast model may have captured some of the dynamics of the hub ensemble. However, more work needs to be done to understand the implicit epidemiological model that represents its behaviour and whether or not this is the optimal choice for stakeholders' requirements. We also conclude that our findings may be largely driven by the choice of evaluation measure used by the forecast hub and that if this is misspecified then the forecast hub submissions may not be optimised to forecast users' expectations or requirements. Our work is potentially useful for forecast consumers as a tool for understanding the inherent assumptions of the forecasts they are making use of and to researchers aiming to develop forecast approaches that perform similarly, or better than, current multi-model and multi-team forecast ensembles that are trusted by stakeholders.
bibliography: library.bib
output:
  bookdown::pdf_book:
    base_format: rticles::peerj_article
    extra_dependencies: ["float"]
---

@@

We evaluate forecasts for all locations and horizons quantitatively using the weighted interval score (WIS) [@bracherEvaluatingEpidemicForecasts2021], which is a quantile-based proper scoring rule that approximates the continuous ranked probability score (CRPS), and the absolute error (AE) of the median forecast.
Both the WIS and CRPS are generalisations of the absolute error to evaluate probabilistic forecasts and are widely used to evaluate COVID-19 forecasts, including by the European forecast hub [@Sherratt2022-ue].
We present WIS for the subset of forecasts we explore visually for both the ensemble and surrogate model by date and forecast horizon (1 and 4 weeks).

To understand the relative performance of the surrogate model compared to the ensemble model, we calculate the relative performance (rWIS and rAE) by dividing the WIS/AE for the surrogate model by the WIS/AE of the ensemble model for all locations and forecast horizons.
To maintain the propriety of this score, we do this after first taking the means of scores for the relevant stratification.
We explore relative performance by forecast horizon, by month and horizon, and by location and horizon.

In addition to presenting the WIS for a subset of locations and the relative WIS for all locations, we also calculate and visualise the empirical coverage, which is the percentage of observed values within a given interval or below a given quantile, of both the surrogate and ensemble model for the 30%, 60%, and 90% prediction intervals and by quantile. 
Lastly, we calculate and visualise the relative weighted interval score by quantile, stratified by forecast horizon, to assess the relative difference in performance across the predictive distribution.

@@

In our example set of locations, the absolute performance of the ensemble and the surrogate model was visually similar on the log scale in all locations at short forecast horizons though this varied by location (Figure \@ref(fig:vis-forecasts) b).
On the natural scale the difference in performance was more marked, specially for periods of high incidence and at longer horizons (Figure \@ref(fig:vis-forecasts) a).
Performance was not homogeneous across our set of example locations with the surrogate model performing similarly to the ensemble in Slovakia whilst in the United Kingdom and Germany the surrogate model performed visibly substantially worse for some forecast dates (Figure \@ref(fig:vis-forecasts)).
For both the ensemble and the surrogate, performance visibly decreased as the forecast horizon increased with this being particularly noticable for the surrogate model during periods of high incidence.
In general, in the study period, the ensemble appeared to be more liable to underpredict and the surrogate model appeared more likely to over-predict. 
Both models forecast large reductions in incidence in Poland during May that did not occur whilst only the ensemble forecast spuriously forecast similar large reductions in Germany during June.
In comparison to the ensemble model the surrogate model appeared less likely to place weight on unfeasibly large reductions in incidence during periods of declining incidence but on other hand was more likely to forecast continuing increases in incidence (for example in February in Slovakia and Poland).

@@

Over the 6 months of the study period, we found that our surrogate model produced forecasts that were visually similar to those from the forecast hub ensemble on the log scale though with greater uncertainty.
Visual differences were more marked on the natural scale with the surrogate model forecasting spuriously high peak incidence.
In a subset of example locations, we observed some variation in performance across locations, that the ensemble appeared biased toward underprediction, and the surrogate model appeared slightly biased toward overprediction.
Evaluating the relative performance of the surrogate model compared to the European forecast hub ensemble we found that the mean performance was substantially worse, which was driven by a relatively small subset of vey poorly performing forecasts.
Median forecast performance of the surrogate model was also worse when compared to forecasts from the ensemble though the majority of surrogate forecasts were within 50% of the performance observed for the ensemble forecast.
Performance varied by location and forecast date with the surrogate model performing worse in the first part of 2022 which may have been linked to incidence rates peaking across forecast locations linked to the spread of BA.2.
In general, the relative performance of the surrogate model degraded as forecast horizons increas with the distribution of relative performance having an increasingly heavy right tail as the forecast horizon increased indicating a greater share of forecasts performing very poorly in comparison to the hub ensemble.
The forecast hub ensemble was poorly calibrated, particularly at longer forecast horizons and bigger prediction intervals, compared to the surrogate model though the surrogate model tended to be overly uncertain at large intervals.
Evaluating point forecast performance indicated a similar pattern of performance as that observed using the full predictive distribution though relative performance of the surrogate model generally improved.

@@

Whilst we derived our surrogate model from a range of assumptions based on observing ensemble forecasts behaviour and the behaviour and structure of submitted models this was by no means as structured or complete a process as it could have been.
In follow-up work, a more rigorous approach to this could be taken to further refine this set of assumptions, in particular using the input of a wider pool of researchers. 
The findings from our study may also be useful for informing this improved set of assumptions. 
A particular focus should be understanding why our surrogate mdoel was liable to overestimate peak incidence and what simply additional assumptions may be used to mitigate this.
In addition, the model we derived based on our assumptions was likely not optimal both in terms of compute time and accuracy at reproducing ensemble-like behaviour.
Models with a more complex auto-correlation structure and more refined approaches to localised trends should be explored to improve relative performance to ensemble forecasts.
An example of a family of possible approaches are structural time series models which have many of the characteristics implied by our assumptions for how forecast ensembles typically operate.
As we identified that the tails of our predictive distributions were responsible for a large proportion of the difference in performance compared to the forecast ensemble it may be the case that post-processing of forecasts from our surrogate model would enhance their similarity to the forecast ensemble.
This seems likely to improve out-of-sample performance but does not help with the secondary aim of understanding the implicit assumptions driving the performance of multi-model, multi-team infectious disease forecast ensembles.
As we have hypothesised that the use of absolute scoring measures is inappropriate and leads to performance characteristics that are unlikely to be favoured by forecast stakeholders more work should be done in this area.
If new forecast ensemble methods are adopted as best practice by forecast hubs then follow-up work attempting to create surrogate forecast models should also use these approaches and this will likely alter the observed characteristics of the hub ensemble forecasts, for example, the tendency to be poorly calibrated.
In September 2022, GitHub announced support for hosted GitHub Action runners with additional compute power [@GitHubActionsLargeRunners]. Whilst a paid feature this may allow more compute attensive models, with fewer potental performance trade-offs, to be easily democratised though only if funds are available to support the hosting costs.
One potential research area is to explore forecasting methods that can be used with a range of compute resources though this would require extensive evaluation and documentation to make it clear to users what the trade-offs between compute usage and forecast performance are.
More work is needed to understand the best practice treatment of data revisions when evaluating forecasts and the potential sources for bias these may cause.
Lastly, here we have only explored a surrogate for an ensemble for a single disease, a limited set of locations, and a single target (incident cases), meaning our findings are difficult to generalise.
Follow-up work should explore whether this behaviour holds across diseases, locations, and epidemiological targets where the behaviour of ensembles is notably different [@REF].
However, this work is limited by the lack of large-scale forecast ensembling projects despite them showing obvious promise to improve the forecasts available to stakeholders.
